---
title: "Lab 1"
author: "Drake Johnson"
date: "10/31/23"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

Submissions should be make in blackboard and consist of exactly 2 files: the R Markdown (rmd) file and the PDF file.

## Problem 1: Inputting data

Find a small dataset with two variables and input columns of values for both variables. (5 Points)
```{r}
time<-c(4,5,5,6,4,7,4,5,5,4)
speed<-c(10,9,9,8,7,11,10,12,9,10)
```

## Problem 2: Calculating statistics
Calculate the mean, standard deviation and Pearson correlation coefficient for your dataset. (10 Points)

```{r}
mean(time)
mean(speed)
sd(time)
sd(speed)
cor(time,speed)
```

## Problem 3: Plot a scatterplot
Create a scatterplot of your dataset. (5 Points)

```{r}
plot(time,speed)
```

## Problem 4: Read in a csv and get a data summary
Find a dataset(.csv), read it and get a summary of the dataset. (10 Points)

```{r}
data<-read.csv("HollywoodsMostProfitableStories.csv")
summary(data)
```

Use the dataset from Problem 4 to asnwer the next problems.

## Problem 5: Create a histogram 

Select a continuous variable in your dataset and create a histogram for that variable. Add chart title, labels, and color. (10 Points)

```{r}
hist(data$Profitability,main = "Histogram of profitability",xlab = "profitability",col = "red",border="green")
```
## Problem 6: Create a Bar chart 

Select a discrete variable in your dataset and create a Bar chart for that variable. Add chart title, labels, and color.  (10 Points)

```{r}
barplot(table(data$Year),main = "bar chart of years",ylab = "movies",col = topo.colors(8))
```

## Problem 7: Create a scatterplot 

Select a two continuous variables in your dataset and create a scatterplot for these variables. Add chart title, labels, and color. (10 Points)

```{r}
plot(data$Audience..score..,data$Rotten.Tomatoes..,main = "audience score vs. rotten tomatoes",xlab = "audience score",ylab = "rotten tomatoes",col="green")
```

## Problem 8: Creating a Dot plot for two variables (Discrete X, Continuous Y) 

Select a two variables (Discrete X, Continuous Y) in your dataset and create a a dot plot for these variables. Add chart title, labels, and color. (10 Points)

```{r}
dotchart(data$Worldwide.Gross,main="Dot plot of worldwide gross",color = heat.colors(8))
```

## Problem 9: Creating a Box plot for two variables (Discrete X, Continuous Y) 

Select a two variables (Discrete X, Continuous Y) in your dataset from problem 4 and create a a Box plot for these variables.Add chart title, labels, and color. (10 Points)
```{r}
boxplot(Profitability~Year, data = data, color=topo.colors(4))
```
8




